---
title: "Problem Set 7"
author: "Rebecca Xi"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(skimr)
library(broom)
library(gt)
library(tidyverse)

# Read in the data

data <- read.csv('raw-data/HSall_members.csv')
```


## Question 1: Conduct exploratory data analysis

### 1A) Summary Statistics

```{r 1A, echo=FALSE}

# I create a tibble from the csv data, filtering for Senators (chamber ==
# “Senate”) and Representatives (chamber == “House”) in the 116th Congress and
# filtering out all independents (party_code == 328). I then create three new
# variables: "party", which records either “Democrat” (party_code == 100) or
# “Republican” (part_code == 200) using an efficient ifelse condition; "age",
# which is 2019 (when the 116th Congress was sworn in) minus birth year; and
# "nominate_percentile", which gives the percentile rank of nominate_dim1 using
# the percent_rank() function. I multiply by 100 to get, say, a value of 50 (for
# 50th percentile) rather than a value of 0.50.

sen_rep_116 <- data %>% 
  filter(
    congress == 116, 
    chamber == "Senate" | chamber == "House",
    party_code != 328
  ) %>% 
  mutate(
    party = ifelse(party_code == 100, "Democrat", "Republican"),
    age = 2019 - born,
    nominate_percentile = 100 * percent_rank(nominate_dim1)
  )

# Here I print the output of calling skim on the nominate_percentile and age
# variables, as an alternative to summary

 skim(sen_rep_116$nominate_percentile)
 skim(sen_rep_116$age)
```


### 1B) Visualizing a single variable

```{r 1B, echo=FALSE}

# Here I create a histogram showing the distribution of congressional ages,
# colored in by party (Democrat or Republican). I use the fill argument to
# indicate party, since I am personally interested in seeing the distribution of
# ages across parties (do Democrats or Republicans generally skew older?). I use
# facet_wrap to wrap the data by House and by Senate so I get an age
# distribution for each one, which is helpful in thinking about (1) the contrast
# in age distributions, if any, between House and Senate,  and (2) the
# distribution of Democrats and Republicans across House and Senate.

age_plot <- sen_rep_116 %>% 
  select(age, party, chamber) %>% 
  ggplot(aes(x = age, fill = party)) +
    geom_histogram(bins = 40) +
    facet_wrap(vars(chamber)) +
    labs(
      title = "Distribution of Congressional Ages",
      subtitle = "for Democrats and Republicans at the 116th Congress",
      source = "Voteview: Congressional Roll-Call Votes Database",
      x = "Age",
      y = "Count"
    ) +
    scale_fill_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
    theme_classic()
age_plot
  
# I copy-paste the above code here and modify for nominate_percentile. The goal
# here is to be able to visualize the split between "liberals" and
# "conservatives", which unsurprisingly corresponds to the split between
# Democrats and Republicans. By eyeballing, I observe a slighter more liberal
# House and a slightly more conservative Senate.

nom_perc_plot <- sen_rep_116 %>% 
  select(nominate_percentile, party, chamber) %>% 
  ggplot(aes(x = nominate_percentile, fill = party)) +
    geom_histogram(bins = 40) +
    facet_wrap(vars(chamber)) +
    labs(
      title = "Distribution of Outcomes",
      subtitle = "for Democrats and Republicans at the 116th Congress",
      source = "Voteview: Congressional Roll-Call Votes Database",
      x = "Percentile Rank (DW Nominate)",
      y = "Count"
    ) +
    scale_fill_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
    theme_classic()
nom_perc_plot
```


### 1C) Bivariate Correlations

```{r 1C, echo=FALSE}

# Here I calculate the correlation coefficient ("corr") for age and
# nominate_percentile

corr <- sen_rep_116 %>% 
  summarize(corr = cor(age, nominate_percentile))
corr
```

A negative correlation coefficient indicates a negative relationship (if one 
variable increases, the value of the other decreases), while a positive 
coefficient indicates a positive relationship (the values of both variables move
in the same direction). 


### 1D) Plotting Bivariate relationships

```{r 1D, echo=FALSE, message=FALSE}

# Here I create a scatterplot of the relationship between age and
# nominate_percentile, coloring each point based on each member’s party. I use
# geom_smooth(method = "glm")) to add a black “best fit” line, illustrating the
# relationship between age and DW nominate percentile. I set se = FALSE to
# remove the shading of the confidence interval around the line. I also add two
# more best fit lines (using the additional argument color = party within aes(),
# in order to split into parties for two lines) for the relationship within
# parties, which we see is roughly similar to the best fit line overall.
# Finally, I scale for desired color and to scale the y-axis by 100.

plot <- sen_rep_116 %>% 
  select(age, nominate_percentile, party) %>% 
  ggplot(aes(x = age, y = nominate_percentile)) +
    geom_point(aes(color = party)) +
    geom_smooth(method = "glm", se = FALSE, color = "black") +
    geom_smooth(method = "glm", se = FALSE, aes(color = party)) +
    labs(
      title = "Relationship between Age and DW Nominate Percentile",
      subtitle = "Higher scores represent more conservative positions",
      x = "Age",
      y = "Percentile Rank (DW Nominate)"
    ) +
    scale_color_manual(
      values = c("Democrat" = "blue", "Republican" = "red"),
      name = "Party",
      labels = c("Democrat", "Republican")
    ) +
    theme_classic()

# Note: I set this code chunk to message=FALSE simply to hide a perfectly
# innocuous informative message, that `geom_smooth()` using formula 'y ~ x'

plot
```



## Question 2: Run a single regression

### 2A) Using lm()

```{r 2A, echo=FALSE}

# I first subset my previous dataset sen_rep_116to just include democrats
# (party_code = 100). (I could just as well do this using party == "Democrat".)

dem_116 <- sen_rep_116 %>% 
  filter(party_code == 100)

# Now I run a linear regression of nominate_percentile on age using lm(). I then
# tidy up the data and select for the variables of interest: term, estimate,
# conf.low, and conf.high. Finally, I print this tidied data in the form of a gt
# table.

dem_116_model <- dem_116 %>% 
  lm(nominate_percentile ~ age, data = .) %>% 
  tidy(conf.int = TRUE) %>% 
  select(term, estimate, conf.low, conf.high) %>% 
  gt() %>% 
  tab_header(
    title = "Effect of Age on DW Nominate Percentile",
    subtitle = "Higher percentile suggests more conservative"
  ) %>% 
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "Lower bound",
    conf.high = "Upper bound"
  )
dem_116_model
```


### 2B) Interpreting results

The slope coefficient on age of -0.33 indicates that the average treatment 
effect of increasing the age of a democratic member of Congress by 1 year is 
-0.33. This means that aging one year has the effect, on a democratic member of 
Congress, of decreasing that member's percentile rank for DW nominate scores by 
about 0.33 percentiles. We should beware of interpreting the estimated 
correlation as causal right away. There may be other reasons or even 
*confounding variables* that affect the relationship between age and level of 
conservatism. One such reason may be that a subset of the older democratic 
members of the 116th Congress were influenced by a common, more liberal 
political ideology, perhaps even in their youth, and that democratic ideology 
has become relatively less liberal since.

We calculate a confidence interval around the average treatment effect to 
reflect the uncertainty that comes with estimation. Since the slope coefficient 
in the linear regression is measured with some uncertainty, the estimate of 
potential DW Nominate percentile outcome under the "treatment" of age is also 
measured with some uncertainty, and we need to take that into consideration.


### 2C) Regression and the Rubin Causal Model

The coefficient on military is equivalent to the average treatment effect (of 
undergoing active military duty) on DW Nominate percentile rankings. It serves 
as a way to quantify the difference in potential outcomes, between the potential
outcome when a member does not undergo military duty and the potential outcome 
when a member does go into the military. Randomization is necessary to support 
this interpretation as causal: random assignment is the best way to infer an 
average treatment effect, given a large enough sample size. 



## Question 3: Generalize to many regressions

```{r 3, echo=FALSE}

# Starting with the entire nominate dataset. I filter for democratic members
# (party_code == 100) in either the Senate (chamber == “Senate”) or the House of
# Representatives (chamber == “House”). I then (re)create my age and
# nominate_percentile variables, updating the age variable from 1A to record the
# difference between the year the congress began (starting in 1789, with
# congress being held every 2 years) and the year the member was born. I then
# use group_by(congress) and nest() to create a tibble of tibbles, with two
# columns (congress and data) and 92 rows (for the 92 congresses). Each row of
# the data column is a list with all the observations for a given congress.

dem <- data %>% 
  filter(
    party_code == 100,
    chamber == "Senate" | chamber == "House"
  ) %>% 
  mutate(
    age = (1789 + 2 * congress) - born,
    nominate_percentile = 100 * percent_rank(nominate_dim1)
  ) %>% 
  group_by(congress) %>% 
  nest()

# Now I create a column "mod", using map() to run lm() on the data in each row
# of the dem tibble (specifically, on nominate_percentile and age, as before). I
# have to be sure to map lm to data specifically, rather than on the congress
# variable or something else, or else the function will not be able to find
# nominate_percentile and age. I then create another column "reg_results" which
# uses map() again to tidy the results. Notably, I found
# https://tidyr.tidyverse.org/articles/nest.html very helpful for mapping with
# lm, tidy, and nested tibbles. I then use map_dbl() to pull the estimate for
# age from each congress and store it in coef, as well as the standard error
# which I store in se. Finally, I use these two variables to create an "upper"
# column and a "lower" column for the 95% confidence interval, using the
# standard error method of calculating confidence intervals.

dem <- dem %>% 
  mutate(
    mod = map(data, function(list) lm(nominate_percentile ~ age, data = list)),
    reg_results = map(mod, function(list) tidy(list, conf.int = TRUE)),
    coef = map_dbl(
      reg_results, 
      ~ filter(., term == "age") %>% 
        pull(estimate)
    ),
    se = map_dbl(
      reg_results,
      ~ filter(., term == "age") %>% 
        pull(std.error)
    ),
    upper = coef + 1.96 * se,
    lower = coef - 1.96 * se
  )

# Lastly, I use the tibble to replicate the given plot. The x-axis indicates
# which congress, and the y axis indicates the average treatment effect (coef)
# of age on nominate_percentile for that congress. I use geom_errorbar(aes(x =
# congress, ymin = lower, ymax = upper)) to produce the 95% confidence intervals
# on the estimates. Finally, I color both the points and the error bars blue,
# write appropriate titles and axis labels, and apply theme_classic() to mimic
# the plot in the problem set instructions.

dem_plot <- dem %>% 
  ggplot(aes(x = congress, y = coef)) +
    geom_point(color = "blue") + 
    geom_errorbar(aes(x = congress, ymin = lower, ymax = upper), color = "blue") +
    labs(
      title = "Average Treatment Effect of Age on Conservativism\n Among Democrats Over Time",
      subtitle = "Scores use DW Nominate Dimension 1\n 95% Confidence Interval",
      x = "Congress",
      y = "Estimate"
    ) +
    theme_classic()
dem_plot
```







